---
title: "EDS230/ESM232 Assignment 3: Informal Sensitivity Analysis"
date: "Spring 2023"
output: html_document
---

### instructions 

+ Develop a profit model for your almond yield (you can make this up - think about what the parameters would be)

    * you might assume a baseline profit and then adjust according to the anomaly  

    * there are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function) 

+ Do a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

+ Create a single graph of the results - you can decide what is the most meaningful graph 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("almond_model.R")
library(tidyverse)

```

### step 1: profit model 

1. Develop a profit model for your almond yield (you can make this up - think about what the parameters would be)

```{r load_functions}

## creating profit function here 
## NPV formula - accounting for time value of money 
## assuming time is by the year
## value is going to be yield 
## rate is $/yield

compute_NPV = function(rate=1, value, time, discount=0.12) {
	result = rate*value / (1 + discount)**time
	return(result)
}

# load almond yield function
source("almond_model.R")

# here is what my almond yield function looks like 
almond_model

```

### step 2: informal sensitivity analysis 
 
2. Do a simple informal sensitivity analysis of total almond yield profit using at least 2 parameters

In this step, I use a wrapper function to combine the two functions for almond yield and for calculating NPV, and then take a summary statistic. 

```{r sa}
climdf <- read.csv("../clim.csv", sep=" ")

# format input data 
temp = climdf %>% 
  group_by(month, year) %>% 
  summarize_at(vars(tmin_c), list(mean)) %>% 
  dplyr::filter(month==2)

pre = climdf %>% group_by(month, year) %>% 
  summarize_at(vars(precip), list(sum)) %>% 
  dplyr::filter(month==1)

# set up the sensitivity parameters

# r is the rate in my NPV function
r = rnorm(50, mean=2, sd=0.5)
# int is the intercept in the almond yield equation 
int = rnorm(50, mean=0.28, sd=0.1)

# create data frame with both parameters, nrow = 50, so will run the model 50 times 
parms = data.frame(r, int)

# here is my wrapper function: it will use each row in my 'parms' dataframe and do the following:
# 1) get almond yield, output for each year  
# 2) convert yield to profit for each year 
# 3) get summary statistic - in this example I am taking the total amount of profit made over the entire time period 
wrapper <- function(r, int){
  yield = almond_model(Tn_2=temp$tmin_c, P_1=pre$precip, intercept = int)
  dollars = compute_NPV(value=yield$yields, rate=r, time=seq(1:length(yield$yields)))
  total_profit = sum(dollars)
  
  return(total_profit)
}

# this is similar to doing a for loop, except now I am going to use 'pmap' to run the wrapper 
sa_parms <- pmap(parms, wrapper)

# unlist the 'pmap' results 
profit_results <- data.frame(net_profit=unlist(sa_parms))

summary(profit_results)
```

What if we want to do the same thing but with a for loop? 
```{r looping}

total_profit = vector(length=50)

for(i in 1:nrow(parms)){
  yield = almond_model(Tn_2=temp$tmin_c, P_1=pre$precip, 
                       intercept = parms$int[i])
  dollars = compute_NPV(value=yield$yields, rate=parms$r[i], time=seq(1:length(yield$yields)))
  total_profit[i] = sum(dollars)
}

summary(total_profit)

```

### step 3: visualize results

3. Create a single graph of the results - you can decide what is the most meaningful graph 

Here you can start simple and just explore results, these are all of examples of visualizing the range of the output.

```{r plot_results}

ggplot(profit_results) + geom_histogram(aes(x=net_profit))
ggplot(profit_results) + geom_density(aes(x=net_profit))
ggplot(profit_results) + geom_boxplot(aes(y=net_profit))

```

We can also plot the results and how they relate to the parameters we varied

```{r plot_parms}

# combine parms dataframe with results
# using pivot longer so that plot can use facet_grid 
all_results = cbind(profit_results, parms) %>% 
  pivot_longer(cols=c(r, int))

ggplot(all_results) + geom_point(aes(x=value, y=net_profit)) + facet_wrap('name', scales="free")


```
